Regression-based

### 2023-12-13 Regression.Rmd
Did some descriptiosn for the linear regression model

```{r}
count <- colSums(is.na(cirrhosis))

kable(count)
```
##### Splitting:

we'll split the dataset into training and testing sets. The training set will be used to train our model, while the testing set will help evaluate its performance. We'll use 80% of the data for training and the remaining 20% for testing.

@@

#cirrhosisLRData <- na.omit(cirrhosisLRData)

set.seed(123)  #reproducibility

@@

##### Test out basic model
```{r}
LRModel1 <- lm( #train the insurance model
  formula = N_Days ~ Platelets,
  data = test_data,
  na.action = "na.omit"
)

@@

##### Define the models

@@

##### Calculating RMSE for first model
```{r}
sqrt(mean(residuals(initialModel)^2, na.rm = TRUE))
```

#####Using stepwise regression

@@

#### Trying various algorithms to choose the best features/attributes

##### Backward Selection
- Backward selection is a method of fitting regression models in which the choice of predictive variables is carried out by an automatic procedure.
- step() function is employed to perform backward selection. The fullModel is used as the starting model, and variables are iteratively removed until the model reaches the specification of interceptModel, which only contains the intercept. The data used for this selection process omits any rows with missing values. The direction = backward argument specifies the backward elimination method, and k = 2 determines the penalty for adding a variable. Finally, trace = 0 indicates that the intermediate steps of the selection process won't be printed. The final model after backward selection is stored in ModelBS

summary of backward:
```{r}
ModelBS <- step(
  object = fullModel,
  scope = list(
    lower = interceptModel,
    upper = fullModel
  ),
  data = na.omit(test_data), 
  direction = "backward",
  k = 2,
  trace = 0
)

@@

##### Forward Selection

@@

##### Stepwise

@@

##### Build QQ Plot for the stepwise Train Model

@@

##### Calculate Skewness of Residuals from the Stepwise Training Model

@@

##### Sample Excess Kurtosis
```{r}
kurtosi(residuals(stepwiseTrainModel))
```

##### Built a plot for the model

@@

### PREDICT
```{r}
predictVals <- predict(
  object = stepwiseTrainModel,
  newdata = test_data
)

@@

##### Plot comparing actual charges to predicted charges from the test_data

the blue reference line indicates wher predictions match actual values perfectly.
```{r}
ggplot(
  data = test_data,
  mapping = aes(x = N_Days, y = prediction)
) +
  geom_point() +
  geom_abline( 
    slope = 1,
    intercept = 0,
    color = "blue",
    linetype = "solid",
    linewidth = 1
  ) +
  theme_bw() +
  labs(
    x = "Actual",
    y = "Predicted"
  )
```

>>>>

# Part II: Prediction for the Number of Days until an event (death)

Next, let us address our next sub-goal: predicting the number of days a patient survives. The objective is to use the medical test data help doctors make better decisions. To do this, we will be using linear regression.

## Introduction

Linear regression is a statistical technique to model the relationship between a dependent variable and one or more independent variables. The model tries to draw a straight line that best fits the data points on a graph, representing how changes in the independent variables are associated with changes in the dependent variable. The primary advantage of linear regression is its simplicity and ease of interpretation. However, its main drawbacks are that it assumes a linear relationship between variables, which may not always be true. This type of model can be very sensitive to outliers, which affects the model's accuracy, especially dealing with medical data. Furthermore, unlike the decision tree, linear regression models cannot ignore N/A values.

![[Linear Regression Example (]{.underline}[Source](https://www.spiceworks.com/tech/artificial-intelligence/articles/what-is-linear-regression/))](LinRegEg.png)

## Methodology

Similar to the tree, we will start by splitting the data set into training and testing sets. Note that the data now onluy contains instances where the patient was deceased. The training set will be used to train our model, while the testing set will help evaluate its performance. We'll use 80% of the data for training and the remaining 20% for testing. To allow reproducible code, we have fixed the seed at 380.

@@

set.seed(380)  #reproducibility

@@

### Step 1: Train basic tree

```{r}
LRModel1 <- lm( #train the basic tree
  formula = N_Days ~ Platelets,
  data = test_data,
  na.action = "na.omit"
)

@@

### Step 2: Train basic models

@@

##### Calculating RMSE for first model

```{r}
sqrt(mean(residuals(initialModel)^2, na.rm = TRUE))
```

### Step 3: Testing selection models

@@

Here, we will be trying various algorithms to choose the best features/attributes

#### (a) Backward Selection

-   Backward selection is a method of fitting regression models in which the choice of predictive variables is carried out by an automatic procedure.

-   step() function is employed to perform backward selection. The fullModel is used as the starting model, and variables are iteratively removed until the model reaches the specification of interceptModel, which only contains the intercept. The data used for this selection process omits any rows with missing values. The direction = backward argument specifies the backward elimination method, and k = 2 determines the penalty for adding a variable. Finally, trace = 0 indicates that the intermediate steps of the selection process won't be printed. The final model after backward selection is stored in ModelBS

summary of backward:

```{r}
ModelBS <- step(
  object = fullModel,
  scope = list(
    lower = interceptModel,
    upper = fullModel
  ),
  data = na.omit(test_data), 
  direction = "backward",
  k = 2,
  trace = 0
)

@@

#### (b) Forward Selection

@@

#### (c) Stepwise

@@

## Step 4: Results

### Build QQ Plot for the stepwise Train Model

@@

### Calculate Skewness of Residuals from the Stepwise Training Model

@@

#### Sample Excess Kurtosis

```{r}
kurtosi(residuals(stepwiseTrainModel))
```

#### Built a plot for the model

@@

### Predictions

```{r}
predictVals <- predict(
  object = stepwiseTrainModel,
  newdata = test_data
)

@@

### Plot predictions

the blue reference line indicates where predictions match actual values perfectly.

```{r}
ggplot(
  data = test_data,
  mapping = aes(x = N_Days, y = prediction)
) +
  geom_point() +
  geom_abline( 
    slope = 1,
    intercept = 0,
    color = "blue",
    linetype = "solid",
    linewidth = 1
  ) +
  theme_bw() +
  labs(
    x = "Actual",
    y = "Predicted"
  )
```
